|

### 2021-03-10 supplementary_information_science.Rmd
Update supplementary_information_science.Rmd

We depicted branching ecosystems as a bifurcating network of connected habitat patches [@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014] (**Figure 1**). Habitat patches, or local communities, can be either non-branching or branching river sections with a unit length $l$, which defines the scale of local species interactions. Two parameters determine the geometric properties of simulated branching networks: the number of habitat patches $N_p$ (ecosystem size) and branching probability $P_b$ (ecosystem complexity). Each habitat patch is assigned to be a branching patch (including upstream terminals) with probability $P_b$ or a non-branching patch with probability $1-P_b$. In this framework, an individual branch is a consecutive series of non-branching patches terminated at a branching patch; therefore, the number of habitat patches in branch $q$, $n_{p,q}$, is a realization of a random variable drawn from a geometric distribution $n_{p,q} \sim Ge(P_{b})$ [@teruiMetapopulationStabilityBranching2018] but conditional on $\sum_{q=1}^{N_b}n_{p,q} = N_p$ ($N_b$ is the number of branches in a network). The number of branches $N_b$ is identical to the number of branching habitat patches in a network, or the number of 'success' out of $N_p$ trials. This leads to an assumption that $N_b$ is a realization of a random variable drawn from a binomial distribution as $N_b \sim Binomial(N_p, P_b)$. This framework has two merits. First, it reflects observed patterns of branch length distribution, which is known to follow an exponential distribution (a continuous version of a geometric distribution) [@rodriguez-iturbeFractalRiverBasins2001]. Second, it preserves the fractal nature of branching patterns [@peckhamReformulationHortonLaws1999]. An alternative approach is the use of optimal channel networks [@carraroGenerationApplicationRiver2020]. However, we employed our framework because branching probability $P_b$ is quantifiable in natural river networks, enabling us to compare theoretical predictions with empirical observations.

@@

\begin{equation}
  \mu_{z,k} = \rho \mu_{z,k+1} + \epsilon_{k}
\end{equation}

where $\mu_{z,k}$ is the mean environmental value at longitudinal position $k$ ($k$ is the network distance from the outlet patch; $k = 1$ at the outlet), $\rho$ is the strength of spatial autocorrelation and $\epsilon_k$ is the local environmental noise that follows a normal distribution with a mean of zero and a standard deviation of $\sigma_l$. The parameter $\rho$ can take values of 0-1 with larger values indicating greater spatial autocorrelation. In this study, we set $\rho = 1$ to mimic strong spatial autocorrelation in riverine environments. At confluences, we took a weighted mean of environmental values given the relative size of upstream contributing area of joining tributaries:

\begin{equation}
\begin{aligned}
  \mu_{z,k} &= \omega(\rho \mu_{z,k+1}^R + \epsilon_{k}^R) + (1-\omega)(\rho \mu_{z,k+1}^L + \epsilon_{k}^L) \\
  \omega &= \frac{A_{k+1}^R}{A_{k+1}^R + A_{k+1}^L}
\end{aligned}
\end{equation}

where $R$ and $L$ denote "right" and "left" branches, respectively, and $A_k$ is the number of upstream habitat patches at longitudinal position $k$ (akin to the upstream watershed area; $A_k = 1$ at the most upstream patch). The parameter $\omega$ takes values of 0-1 and represents the relative influence of the right tributary given the relative size of joining tributaries. With this expression, larger tributaries have a greater influence on the downstream environment, as observed in natural river networks [@bendaNetworkDynamicsHypothesis2004].

@@

\begin{equation}
  N_{ix}(t + 1) \sim Poisson(n_{ix}(t) + I_{ix}(t) - E_{ix}(t))
\end{equation}

where $n_{ix}(t)$ is the expected number of individuals given the local community dynamics at time $t$, $I_{ix}(t)$ the expected number of immigrants to patch $x$, and $E_{ix}(t)$ the expected number of emigrants from patch $x$. The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes [@fronhoferClassicalMetapopulationDynamics2017; @thompsonProcessbasedMetacommunityFramework2020]. Local community dynamics are simulated based on the Beverton-Holt equation [@bevertonDynamicsExploitedFish1957]:

\begin{equation}
  n_{ix}(t) = \frac{N_{ix}(t)r_{ix}(t)}{1 + \frac{r_{0,i}-1}{K_{x}}\sum_{j=1}^S{\alpha_{ij}N_{jx}(t)}}
\end{equation}

where $r_{ix}(t)$ is the reproductive rate of species $i$ given the environmental condition at patch $x$ and time $t$, $r_{0,i}$ the maximum reproductive rate of species $i$, $K_{x}$ the carrying capacity at patch $x$, $\alpha_{ij}$ the competition coefficient between species $i$ and $j$, and $S$ the number of species in a metacommunity. $K_x$ was expressed as a function of the number of upstream habitat patches $A_x$ ($K_x = 100A_x$) to represent increasing habitat size with increasing upstream watershed area [@teruiThreeEcologicalFactors2016; @rahelFishAssemblagesHabitat1991]. The parameter $\alpha_{ij}$ represents the strength of interspecific competition relative to that of intraspecific competition: interspecific competition is greater than intraspecific competition if $\alpha_{ij}$ \> 1 (intraspecific competition coefficient was set constant as $\alpha_{ii} = 1$). $\alpha_{ij}$ was drawn randomly from a uniform distribution as $\alpha_{ij} \sim Unif(0, \alpha_{max})$. The density-independent reproductive rate $r_{ix}(t)$ is affected by abiotic environments (non-consumable) and determined by the following Gaussian function:

\begin{equation}
  r_{ix}(t) = \delta~r_{0,i}~exp[{-\frac{\{\mu_i-z_x(t)\}^2}{2\sigma^2_{niche,i}}}]
\end{equation}

where $\mu_i$ is the optimal environmental value for species $i$, $z_x(t)$ the environmental value at patch $x$ and time $t$, and $\sigma_{niche,i}$ the niche width of species $i$. We introduced the cost of a wide niche by multiplying $\delta$ [@chaianunpornEvolutionaryResponsesClimate2015]:

\begin{equation}
  \delta = exp(-\frac{\sigma_{niche,i}^2}{2\nu^2})
\end{equation}

Smaller values of $\nu$ imply greater costs of having a wider niche (i.e., decreased maximum reproductive rate) and reduce the realized reproductive rate $r_{ix}(t)$ near the niche optimum.

@@

\begin{equation}
  \Omega_{xy} = \sigma_z^2~exp(-\phi d_{xy})
\end{equation}

$\Omega_{xy}$ denotes the temporal covariance of environmental conditions between patch $x$ and $y$, which is assumed to decay exponentially with increasing distance between the patches $d_{xy}$ (the shortest network path measured as the number of habitat patches). The parameter $\phi$ determines the degree of the distance decay of environmental correlates.

@@

\begin{equation}
  \xi_{ix}(t) = \frac{\sum_{y,y \neq x}E_{iy}(t)~exp(-\theta d_{xy})}{\sum_{x} \sum_{y,y \neq x}E_{iy}(t)~exp(-\theta d_{xy})}
\end{equation}

The parameter $\theta$ regulates the dispersal distance of a species that follows an exponential distribution ($\theta^{-1}$ corresponds to the expected dispersal distance). The expected number of immigrants is calculated as $I_{ix}(t) = \xi_{ix}(t)\sum_x^{N_p} E_{ix}$.

@@

\begin{equation}
  \mu_w = \xi_0 + \xi_1~log_{10}~WA_w + \xi_2~log_{10}~P_{b,w} + \xi_3~region_w + \sum_k \xi_k x_{k,w}
\end{equation}

where $\xi_k$ are the intercept and regression coefficients for linear predictors. Meanwhile, the region-specific model assumes the effects of watershed area and branching probability differ between regions (i.e., interactions between region and watershed area/branching probability).

\begin{equation}
\begin{aligned}
  \mu_w =& \xi_0 + \xi_1~log_{10}~WA_w + \xi_2~log_{10}~P_{b,w} + \xi_3~region_w +\\
  &\xi_4~log_{10}~WA_w \cdot region_w + \xi_5~log_{10}~P_{b,w} \cdot region_w + \sum_k \xi_k x_{k,w}
\end{aligned}
\end{equation}

All explanatory variables except watershed area, branching probability, and region were standardized to a mean of zero and a standard deviation of one before the analysis.

@@

\begin{equation}
  Bayes~factor = exp(\Delta BIC_{10}/2)
\end{equation}

where $\Delta BIC_{10} = BIC(M1) - BIC(M0)$.

@@

\begin{equation}
\begin{aligned}
  log_{10}~y_j &\sim Normal(\mu_j, \sigma^2)\\
         \mu_j &= \psi_0 + \psi_1~log_{10}~N_{p,j} + \psi_2~log_{10}~P_{b,j}
\end{aligned}
\end{equation}

where $\psi_k$ ($k = 0-2$) are the intercept ($\psi_0$) and regression coefficients ($\psi_1$ and $\psi_2$). We extracted 500 estimates of $\psi_1$ and $\psi_2$, which represent the effects of $N_{p}$ and $P_{b}$ on diversity metrics under a given parameter combination. To examine influences of simulation parameters (**Table S1**) on $\psi_1$ and $\psi_2$, we developed the following regression model taking $\psi_1$ or $\psi_2$ as a response variable $u_n$ (parameter combination $n$):

\begin{equation}
\begin{aligned}
  u_{n} &\sim Normal(\mu_n, \sigma^2)\\
  \mu_n &= \zeta_0 + \zeta_1 \sigma_{h,n} + \zeta_2 \sigma_{l,n} + \zeta_3 \sigma_{z,n} + \zeta_4 \phi_n + \zeta_5 \nu_n + \zeta_6 \alpha_{max, n} + \zeta_7 \theta_n + \zeta_8 p_{d,n} 
\end{aligned}
\end{equation}

where $\zeta_k$ ($k = 0-8$) are the intercept ($\zeta_0$) and regression coefficients ($\zeta_{1-8}$). Explanatory variables were standardized to a mean of zero and a standard deviation of one, so that regression coefficients are comparable.

@@

\begin{equation}
  \%~bias = \frac{100(S_{est} - S_{true})}{S_{true}}
\end{equation}

Positive values of % bias indicate an overestimation of species richness, while negative values indicate an underestimation.

@@

For regression models, standardized regression coefficients (or its variant) are perhaps the most common summary when comparing effect sizes of explanatory variables. These values are useful if they are directly interpretable. For example, in linear regression models without variable transformations, regression coefficients have direct interpretations as they represent additive effects. However, there are many cases where regression coefficients are difficult to interpret. In our case, we regressed the log-transformed species richness on explanatory variables, so regression coefficients do not have direct interpretations at the original scale of the response variable $y$ (i.e., species richness).

@@

\begin{equation}
  \delta_u(u^{(1)} \to u^{(2)}, v, \Theta) = \frac{E(y|u^{(2)}, v, \Theta) - E(y|u^{(1)}, v, \Theta)}{u^{(2)} - u^{(1)}}
\end{equation}

where $E(\cdot)$ is a known function (e.g., exponential), $u$ the input of interest (a value for the explanatory variable of interest), $v$ all the other inputs (a vector in general), $\Theta$ a set of parameters in a regression model. In general, $\delta_u$ depends on $u^{(1)}$ and $u^{(2)}$ (the lower and higher points of the hypothesized change in the explanatory variable of interest), $v$, and $\Theta$. Therefore, Gelman and Pardoe [@gelmanAveragePredictiveComparisons2007] defined the *average predictive comparison* $\Delta_u$, the mean value of $\delta_u$ over specified distributions of the inputs and model parameters:

\begin{equation}
  \Delta_u = \frac{\int\int_{u^{(2)}>u^{(1)}}du^{(1)}du^{(2)}\int dv \int d\Theta (E(y|u^{(2)}, v, \Theta) - E(y|u^{(1)}, v, \Theta))~p(u^{u^{(1)}}|v)~p(u^{u^{(2)}}|v)~p(v)~p(\Theta)}
                  { \int\int_{u^{(2)}>u^{(1)}}du^{(1)}du^{(2)}\int dv \int d\Theta~(u^{(2)} - u^{(1)})~ p(u^{u^{(1)}}|v)~p(u^{u^{(2)}}|v)~p(v)~p(\Theta) }
\end{equation}

However, directly using the above equation is impractical because estimating $p(u^{(1)}|v)$ and $p(u^{(2)}|v)$ from finite data points is challenging (especially when $v$ is continuous). Instead, we estimated $\Delta_u$ using the following equation proposed by Gelman and Pardoe [@gelmanAveragePredictiveComparisons2007]:

\begin{equation}
  \hat\Delta_u = \frac{\sum_{i=1}^n \sum_{j=1}^n \sum_{s=1}^S w_{ij}(E(y|u_j, v_i, \Theta_s) - E(y|u_i, v_i, \Theta_s))sign(u_j - u_i)}
                      {\sum_{i=1}^n \sum_{j=1}^n \sum_{s=1}^S w_{ij}(u_j - u_i)sign(u_j - u_i)}
\end{equation}

The summations over $n$ data ($i$ and $j$ are a given data point) and $S$ parameter replicates ($s$ is a given simulation replicate) are a realization of averaging over the distributions of $(u^{(1)},v)$ , $u^{(2)}$, and $\Theta$. The factor $w_{ij}$ is a weight that serves to approximate $p(u^{(1)}|v)$ and $p(u^{(2)}|v)$. In theory, $v$ must be held constant while the input of interest changes from $u^{(1)}$ to $u^{(2)}$. However, there are, if any, few transitions from $u^{(1)}$ to $u^{(2)}$ with a common $v$. We approximate such exact transitions by assigning each pair of data points with a weight:

\begin{equation}
  w_{ij} = w(v_i, v_j)
\end{equation}

which should represent the likelihood of $u$ changing from $u_i$ to $u_j$ with $v_i$ (note that $v_i$ is in general a vector of explanatory variables for data point $i$). We used the following weighting function based on Mahalanobis distances following Gelman and Pardoe [@gelmanAveragePredictiveComparisons2007]:

\begin{equation}
w(v_i, v_j) = \frac{1}{1 + (v_i - v_j)^T\sum_v^{-1}(v_i-v_j)}
\end{equation}

where $\sum_v$ is the variance-covariance matrix of $v$.

@@

\begin{equation}
  \hat \Delta_{u,s} = \frac{\sum_{i=1}^n \sum_{j=1}^n w_{ij}(E(y|u_j, v_i, \Theta_s) - E(y|u_i, v_i, \Theta_s))sign(u_j - u_i)}
                           {\sum_{i=1}^n \sum_{j=1}^n w_{ij}(u_j - u_i)sign(u_j - u_i)}
\end{equation}

Then

\begin{equation}
  S.E.(\hat \Delta_{u,s}) = \sqrt{\frac{1}{S-1} \sum_{s=1}^S (\hat \Delta_{u,s} - \hat \Delta_{u})^2}
\end{equation}

The estimated average predictive comparisons were summarized in **Table S7**.

>>>>

We depicted branching ecosystems as a bifurcating network of connected habitat patches [@teruiMetapopulationStabilityBranching2018; @yeakelSynchronisationStabilityRiver2014] (**Figure 1**). Habitat patches, or local communities, can be either non-branching or branching river sections with a unit length $l$, which defines the scale of local species interactions. Two parameters determine the geometric properties of simulated branching networks: the number of habitat patches $N_p$ (ecosystem size) and branching probability $P_b$ (ecosystem complexity). Each habitat patch is assigned to be a branching patch (including upstream terminals) with probability $P_b$ or a non-branching patch with probability $1-P_b$. In this framework, an individual branch is a consecutive series of non-branching patches terminated at a branching patch; therefore, the number of habitat patches in branch $q$, $n_{p,q}$, is a realization of a random variable drawn from a geometric distribution $n_{p,q} \sim Ge(P_{b})$ [@teruiMetapopulationStabilityBranching2018] but conditional on $\sum_{q=1}^{N_b}n_{p,q} = N_p$ ($N_b$ is the number of branches in a network). The number of branches $N_b$ is identical to the number of branching habitat patches in a network, or the number of 'success' out of $N_p$ trials. This leads to an assumption that $N_b$ is a realization of a random variable drawn from a binomial distribution as $N_b \sim Binomial(N_p, P_b)$. This framework has two merits. First, it reflects observed patterns of branch length distribution, which is known to follow an exponential distribution (a continuous version of a geometric distribution) [@rodriguez-iturbeFractalRiverBasins2001]. Second, it preserves the fractal nature of branching patterns [@peckhamReformulationHortonLaws1999]. An alternative approach is the use of optimal channel networks [@carraroGenerationApplicationRiver2020a]. However, we employed our framework because branching probability $P_b$ is quantifiable in natural river networks, enabling us to compare theoretical predictions with empirical observations.

@@

```{=tex}
\begin{equation}
  \mu_{z,k} = \rho \mu_{z,k+1} + \epsilon_{k}
\end{equation}
```
where $\mu_{z,k}$ is the mean environmental value at longitudinal position $k$ ($k$ is the network distance from the outlet patch; $k = 1$ at the outlet), $\rho$ is the strength of spatial autocorrelation and $\epsilon_k$ is the local environmental noise that follows a normal distribution with a mean of zero and a standard deviation of $\sigma_l$. The parameter $\rho$ can take values of 0-1 with larger values indicating greater spatial autocorrelation. In this study, we set $\rho = 1$ to mimic strong spatial autocorrelation in riverine environments. At confluences, we took a weighted mean of environmental values given the relative size of upstream contributing area of joining tributaries:

```{=tex}
\begin{equation}
\begin{aligned}
  \mu_{z,k} &= \omega(\rho \mu_{z,k+1}^R + \epsilon_{k}^R) + (1-\omega)(\rho \mu_{z,k+1}^L + \epsilon_{k}^L) \\
  \omega &= \frac{A_{k+1}^R}{A_{k+1}^R + A_{k+1}^L}
\end{aligned}
\end{equation}
```
where $R$ and $L$ denote "right" and "left" branches, respectively, and $A_k$ is the number of upstream habitat patches at longitudinal position $k$ (akin to the upstream watershed area; $A_k = 1$ at the most upstream patch). The parameter $\omega$ takes values of 0-1 and represents the relative influence of the right tributary given the relative size of joining tributaries. With this expression, larger tributaries have a greater influence on the downstream environment, as observed in natural river networks [@bendaNetworkDynamicsHypothesis2004].

@@

```{=tex}
\begin{equation}
  N_{ix}(t + 1) \sim Poisson(n_{ix}(t) + I_{ix}(t) - E_{ix}(t))
\end{equation}
```
where $n_{ix}(t)$ is the expected number of individuals given the local community dynamics at time $t$, $I_{ix}(t)$ the expected number of immigrants to patch $x$, and $E_{ix}(t)$ the expected number of emigrants from patch $x$. The realized discrete number of individuals is drawn from a Poisson distribution to account for stochasticity in demographic and dispersal processes [@fronhoferClassicalMetapopulationDynamics2017; @thompsonProcessbasedMetacommunityFramework2020]. Local community dynamics are simulated based on the Beverton-Holt equation [@bevertonDynamicsExploitedFish1957]:

```{=tex}
\begin{equation}
  n_{ix}(t) = \frac{N_{ix}(t)r_{ix}(t)}{1 + \frac{r_{0,i}-1}{K_{x}}\sum_{j=1}^S{\alpha_{ij}N_{jx}(t)}}
\end{equation}
```
where $r_{ix}(t)$ is the reproductive rate of species $i$ given the environmental condition at patch $x$ and time $t$, $r_{0,i}$ the maximum reproductive rate of species $i$, $K_{x}$ the carrying capacity at patch $x$, $\alpha_{ij}$ the competition coefficient between species $i$ and $j$, and $S$ the number of species in a metacommunity. $K_x$ was expressed as a function of the number of upstream habitat patches $A_x$ ($K_x = 100A_x$) to represent increasing habitat size with increasing upstream watershed area [@teruiThreeEcologicalFactors2016; @rahelFishAssemblagesHabitat1991]. The parameter $\alpha_{ij}$ represents the strength of interspecific competition relative to that of intraspecific competition: interspecific competition is greater than intraspecific competition if $\alpha_{ij}$ \> 1 (intraspecific competition coefficient was set constant as $\alpha_{ii} = 1$). $\alpha_{ij}$ was drawn randomly from a uniform distribution as $\alpha_{ij} \sim Unif(0, \alpha_{max})$. The density-independent reproductive rate $r_{ix}(t)$ is affected by abiotic environments (non-consumable) and determined by the following Gaussian function:

```{=tex}
\begin{equation}
  r_{ix}(t) = \delta~r_{0,i}~exp[{-\frac{\{\mu_i-z_x(t)\}^2}{2\sigma^2_{niche,i}}}]
\end{equation}
```
where $\mu_i$ is the optimal environmental value for species $i$, $z_x(t)$ the environmental value at patch $x$ and time $t$, and $\sigma_{niche,i}$ the niche width of species $i$. We introduced the cost of a wide niche by multiplying $\delta$ [@chaianunpornEvolutionaryResponsesClimate2015]:

```{=tex}
\begin{equation}
  \delta = exp(-\frac{\sigma_{niche,i}^2}{2\nu^2})
\end{equation}
```
Smaller values of $\nu$ imply greater costs of having a wider niche (i.e., decreased maximum reproductive rate) and reduce the realized reproductive rate $r_{ix}(t)$ near the niche optimum.

@@

```{=tex}
\begin{equation}
  \Omega_{xy} = \sigma_z^2~exp(-\phi d_{xy})
\end{equation}
```
$\Omega_{xy}$ denotes the temporal covariance of environmental conditions between patch $x$ and $y$, which is assumed to decay exponentially with increasing distance between the patches $d_{xy}$ (the shortest network path measured as the number of habitat patches). The parameter $\phi$ determines the degree of the distance decay of environmental correlates.

@@

```{=tex}
\begin{equation}
  \xi_{ix}(t) = \frac{\sum_{y,y \neq x}E_{iy}(t)~exp(-\theta d_{xy})}{\sum_{x} \sum_{y,y \neq x}E_{iy}(t)~exp(-\theta d_{xy})}
\end{equation}
```
The parameter $\theta$ regulates the dispersal distance of a species that follows an exponential distribution ($\theta^{-1}$ corresponds to the expected dispersal distance). The expected number of immigrants is calculated as $I_{ix}(t) = \xi_{ix}(t)\sum_x^{N_p} E_{ix}$.

@@

```{=tex}
\begin{equation}
  \mu_w = \xi_0 + \xi_1~log_{10}~WA_w + \xi_2~log_{10}~P_{b,w} + \xi_3~region_w + \sum_k \xi_k x_{k,w}
\end{equation}
```
where $\xi_k$ are the intercept and regression coefficients for linear predictors. Meanwhile, the region-specific model assumes the effects of watershed area and branching probability differ between regions (i.e., interactions between region and watershed area/branching probability).

```{=tex}
\begin{equation}
\begin{aligned}
  \mu_w =& \xi_0 + \xi_1~log_{10}~WA_w + \xi_2~log_{10}~P_{b,w} + \xi_3~region_w +\\
  &\xi_4~log_{10}~WA_w \cdot region_w + \xi_5~log_{10}~P_{b,w} \cdot region_w + \sum_k \xi_k x_{k,w}
\end{aligned}
\end{equation}
```
All explanatory variables except watershed area, branching probability, and region were standardized to a mean of zero and a standard deviation of one before the analysis.

@@

```{=tex}
\begin{equation}
  Bayes~factor = exp(\Delta BIC_{10}/2)
\end{equation}
```
where $\Delta BIC_{10} = BIC(M1) - BIC(M0)$.

@@

```{=tex}
\begin{equation}
\begin{aligned}
  log_{10}~y_j &\sim Normal(\mu_j, \sigma^2)\\
         \mu_j &= \psi_0 + \psi_1~log_{10}~N_{p,j} + \psi_2~log_{10}~P_{b,j}
\end{aligned}
\end{equation}
```
where $\psi_k$ ($k = 0-2$) are the intercept ($\psi_0$) and regression coefficients ($\psi_1$ and $\psi_2$). We extracted 500 estimates of $\psi_1$ and $\psi_2$, which represent the effects of $N_{p}$ and $P_{b}$ on diversity metrics under a given parameter combination. To examine influences of simulation parameters (**Table S1**) on $\psi_1$ and $\psi_2$, we developed the following regression model taking $\psi_1$ or $\psi_2$ as a response variable $u_n$ (parameter combination $n$):

```{=tex}
\begin{equation}
\begin{aligned}
  u_{n} &\sim Normal(\mu_n, \sigma^2)\\
  \mu_n &= \zeta_0 + \zeta_1 \sigma_{h,n} + \zeta_2 \sigma_{l,n} + \zeta_3 \sigma_{z,n} + \zeta_4 \phi_n + \zeta_5 \nu_n + \zeta_6 \alpha_{max, n} + \zeta_7 \theta_n + \zeta_8 p_{d,n} 
\end{aligned}
\end{equation}
```
where $\zeta_k$ ($k = 0-8$) are the intercept ($\zeta_0$) and regression coefficients ($\zeta_{1-8}$). Explanatory variables were standardized to a mean of zero and a standard deviation of one, so that regression coefficients are comparable.

@@

```{=tex}
\begin{equation}
  \%~bias = \frac{100(S_{est} - S_{true})}{S_{true}}
\end{equation}
```
Positive values of % bias indicate an overestimation of species richness, while negative values indicate an underestimation.

@@

For regression models, standardized regression coefficients (or its variant) are perhaps the most common summary when comparing effect sizes of explanatory variables. These values are useful if they are directly interpretable. For example, in linear regression models without interactions and/or variable transformations, regression coefficients have direct interpretations as they represent additive effects. However, there are many cases where regression coefficients are difficult to interpret. In our case, we regressed the log-transformed species richness on explanatory variables, so regression coefficients do not have direct interpretations at the original scale of the response variable $y$ (i.e., species richness).

@@

```{=tex}
\begin{equation}
  \delta_u(u^{(1)} \to u^{(2)}, v, \Theta) = \frac{E(y|u^{(2)}, v, \Theta) - E(y|u^{(1)}, v, \Theta)}{u^{(2)} - u^{(1)}}
\end{equation}
```
where $E(\cdot)$ is a known function (e.g., exponential), $u$ the input of interest (a value for the explanatory variable of interest), $v$ all the other inputs (a vector in general), $\Theta$ a set of parameters in a regression model. In general, $\delta_u$ rests on $u^{(1)}$ and $u^{(2)}$ (the lower and higher points of the hypothesized change in the explanatory variable of interest), $v$, and $\Theta$. Therefore, Gelman and Pardoe [@gelmanAveragePredictiveComparisons2007] defined the *average predictive comparison* $\Delta_u$, the mean value of $\delta_u$ over specified distributions of the inputs and model parameters:

```{=tex}
\begin{equation}
  \Delta_u = \frac{\int\int_{u^{(2)}>u^{(1)}}du^{(1)}du^{(2)}\int dv \int d\Theta (E(y|u^{(2)}, v, \Theta) - E(y|u^{(1)}, v, \Theta))~p(u^{u^{(1)}}|v)~p(u^{u^{(2)}}|v)~p(v)~p(\Theta)}
                  { \int\int_{u^{(2)}>u^{(1)}}du^{(1)}du^{(2)}\int dv \int d\Theta~(u^{(2)} - u^{(1)})~ p(u^{u^{(1)}}|v)~p(u^{u^{(2)}}|v)~p(v)~p(\Theta) }
\end{equation}
```
However, directly using the above equation is impractical because estimating $p(u^{(1)}|v)$ and $p(u^{(2)}|v)$ from finite data points is challenging (especially when $v$ is continuous). Instead, we estimated $\Delta_u$ using the following equation proposed by Gelman and Pardoe [@gelmanAveragePredictiveComparisons2007]:

```{=tex}
\begin{equation}
  \hat\Delta_u = \frac{\sum_{i=1}^n \sum_{j=1}^n \sum_{s=1}^S w_{ij}(E(y|u_j, v_i, \Theta_s) - E(y|u_i, v_i, \Theta_s))sign(u_j - u_i)}
                      {\sum_{i=1}^n \sum_{j=1}^n \sum_{s=1}^S w_{ij}(u_j - u_i)sign(u_j - u_i)}
\end{equation}
```
The summations over $n$ data ($i$ and $j$ are a given data point) and $S$ parameter replicates ($s$ is a given simulation replicate) are a realization of averaging over the distributions of $(u^{(1)},v)$ , $u^{(2)}$, and $\Theta$. The factor $w_{ij}$ is a weight that serves to approximate $p(u^{(1)}|v)$ and $p(u^{(2)}|v)$. In theory, $v$ must be held constant while the input of interest changes from $u^{(1)}$ to $u^{(2)}$. However, there are, if any, few transitions from $u^{(1)}$ to $u^{(2)}$ with a common $v$. We approximate such exact transitions by assigning each pair of data points with a weight:

```{=tex}
\begin{equation}
  w_{ij} = w(v_i, v_j)
\end{equation}
```
The weight factor should represent the likelihood of $u$ changing from $u_i$ to $u_j$ with a common $v = v_i$ (note that $v_i$ is in general a vector of explanatory variables for data point $i$). We used the following weighting function based on Mahalanobis distances following Gelman and Pardoe [@gelmanAveragePredictiveComparisons2007]:

```{=tex}
\begin{equation}
w(v_i, v_j) = \frac{1}{1 + (v_i - v_j)^T\Omega_v^{-1}(v_i-v_j)}
\end{equation}
```
where $\Omega_v^{-1}$ is the inverse of the variance-covariance matrix of $v$.

@@

```{=tex}
\begin{equation}
  \hat \Delta_{u,s} = \frac{\sum_{i=1}^n \sum_{j=1}^n w_{ij}(E(y|u_j, v_i, \Theta_s) - E(y|u_i, v_i, \Theta_s))sign(u_j - u_i)}
                           {\sum_{i=1}^n \sum_{j=1}^n w_{ij}(10^{u_j} - 10^{u_i})sign(u_j - u_i)}
\end{equation}
```
We exponentiated $u_i$ and $u_j$ because we log-transformed watershed area and branching probability in the statistical model. Then

```{=tex}
\begin{equation}
  S.E.(\hat \Delta_{u,s}) = \sqrt{\frac{1}{S-1} \sum_{s=1}^S (\hat \Delta_{u,s} - \hat \Delta_{u})^2}
\end{equation}
```
The estimated average predictive comparisons were summarized in **Table S7**.
